Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-09-29 psy_peds_3.Rmd
graphics sorted and filtered by week only

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
```

@@

```{r}
# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered_age %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(clinical_table, by = c("patient_num", "days_since_admission")) %>% 
   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
```

@@

# Filtering hospitalizations where patient age is in inclusion criteria
observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

@@

clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
            by = c("patient_num", "days_since_admission")) %>% 
  select(patient_num, n_hospitalisation, age_time_diagnosis)

clinical_table_ado <- clinical_table %>% 
  inner_join(clinical_test[, c("patient_num", "n_hospitalisation", "age_time_diagnosis")],
             by = c("patient_num", "n_hospitalisation"))

@@

all_codes_ado <- all_codes %>% 
   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

head(observation_table_ado)
head(clinical_table_ado)
head(all_codes_ado)
```

@@

### Total hospitalize patients

```{r, fig.height=12, fig.width=12}
total_hospitalizations_per_day <- clinical_filtered %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date ) %>%
  summarise(count = n_distinct(patient_num))

unique_patients_hospitalized_per_week <- clinical_filtered %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count = n_distinct(patient_num))

unique_patients_hospitalized_per_month <- clinical_filtered %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month ) %>%
  summarise(count = n_distinct(patient_num))

th_day <- ggplot(total_hospitalizations_per_day, aes(x = date,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per day)")

th_week <- ggplot(unique_patients_hospitalized_per_week, aes(x = weeks,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per week)")

th_month <- ggplot(unique_patients_hospitalized_per_month, aes(x = month,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per month)")

require(gridExtra)
grid.arrange(th_day, th_week, th_month, ncol=1)
```

#### Group by sex

#### Per week

```{r}
# counts per week and sex
patients_hospitalized_per_week_agg_sex <- clinical_filtered %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( weeks, sex ) %>%
  summarise(count_sex = n_distinct(patient_num))

# the total counts per week was estimated before: unique_patients_hospitalized_per_week
patients_hospitalized_per_week_agg_sex_perc <- unique_patients_hospitalized_per_week %>%
  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
  replace_na(list(count_sex = 0))

patients_hospitalized_per_week_agg_sex_perc %>%
  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage",
       x = "Week",
       title = "Weekly patient (%) aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

## Second estimation: total number of patients filtering by psychiatric disorders

@@

all_codes_desc <- all_codes %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, sex, age_time_visit, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

@@

@@

patients_hospitalized_per_week_agg_sex_perc %>%
  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage ",
       x = "Week",
       title = "Weekly patient (%) aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

#### Psychiatric related conditions vs. any other condition

##### New admissions

##### Total hospitalizations

We have estimated the total number of hospitalizations in the previous
steps: - total_hospitalization_per_day - total_hospitalization_per_week
- total_hospitalization_per_month

We will use these counts to estimate the percentage and to compare total
counts, vs psychiatric related conditions.

```{r}
count_icd_per_day_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd_per_day <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd_per_day <- count_icd_per_day %>%
  left_join(count_icd_per_day_psy, 
            by = c("date", "period")) %>% 
  replace_na(list(count_psy = 0))

count_icd_per_day$percentage_psy <- 100*count_icd_per_day$count_psy / count_icd_per_day$count_icd

thpd_psyPer <- count_icd_per_day %>%
  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer

count_icd_per_day %>%
  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_point()
  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
```

By week

thpd_psyPer <- count_icd_per_week %>%
  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by week)",
       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer

count_icd_per_week %>%
  ggplot(aes(x = weeks, y = percentage_psy, fill = period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by week)",
       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
  geom_point()
  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
```

# Per day

```{r}
thpd_psyCount <- count_icd_per_day %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by day)",
       title = "Daily patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount

# Per week

```{r}
thpd_psyCount <- count_icd_per_week %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by week)",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount

We will compare the psychiatric conditions with some other disorders
(e.g, digestive, respiratory, etc.) We will use these counts to estimate
the percentage and to compare total counts, vs psychiatric related
conditions.

```{r}
total_hospitalizations_per_day_disorder_group <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
        icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system" ))%>%
    group_by( date, icd_description_category ) %>%
  summarise(counts = n_distinct(patient_num))

thpd_disorderGroup <- total_hospitalizations_per_day_disorder_group %>%
  ggplot(aes(x = date, y = counts, fill = icd_description_category, color = icd_description_category)) +
  #geom_point(alpha=0.2) +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "patient number",
       x = "days",
       title = "Total number of hospitalized patients (daily)") +   
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_disorderGroup
```

##### New admissions

##### Total hospitalizations

We estimate and plot the counts per disorder group.

Per day

```{r}
count_disorder_group_per_day <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date, disorder_group ) %>%
  summarise(count_dg = n_distinct(patient_num)) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg_per_day <- count_disorder_group_per_day %>%  
  ggplot(aes(x = date, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Days",
       y = "Patient Count",
       title = "Daily patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg_per_day
```

Per week

Per month

```{r}
count_disorder_group_per_month <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month, disorder_group ) %>%
  summarise(count_dg = n_distinct(patient_num)) %>%
  mutate( period = ifelse( month < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg_per_month <- count_disorder_group_per_month %>%  
  ggplot(aes(x = month, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Months",
       y = "Patient Count",
       title = "Monthly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg_per_month
```

Percentage using as denominator total number of patients per day with a
psy disorder.

Per day - Total counts previously estimated (count_icd_per_day_psy). -
Counts per disorder group previously estimated
(count_disorder_group_per_day)

```{r}
# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_day <- count_icd_per_day_psy %>%
  left_join(count_disorder_group_per_day, 
            by = c("date", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_psy

dgPerc_per_day <- perc_disorder_group_per_day %>%  
  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Days",
       y = "Patient (%)",
       title = "Percentage of daily patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_day
```

Per week - Counts per disorder group previously estimated
(count_disorder_group_per_week)

```{r}
# estimate the weekly counts
count_icd_per_week_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  mutate( period = ifelse( weeks < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_week <- count_icd_per_week_psy %>%
  left_join(count_disorder_group_per_week, 
            by = c("weeks", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_psy

Months

```{r}
# estimate the monthly counts
count_icd_per_month_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  mutate( period = ifelse( month < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_month <- count_icd_per_month_psy %>%
  left_join(count_disorder_group_per_month, 
            by = c("month", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_psy

dgPerc_per_month <- perc_disorder_group_per_month %>%  
  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Months",
       y = "Patient (%)",
       title = "Percentage of monthly patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_month
```

Per day - Counts per disorder group previously estimated
(count_disorder_group_per_day)

```{r}
# estimate the total number per day 
count_icd_per_day <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( date ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_day <- count_icd_per_day %>%
  left_join(count_disorder_group_per_day, 
            by = c("date", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_icd

dgPerc_per_day <- perc_disorder_group_per_day %>%  
  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Days",
       y = "Patient (%)",
       title = "Percentage of daily patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_day
```

Per week - Counts per disorder group previously estimated
(count_disorder_group_per_week)

Months

```{r}
# estimate the monthly counts
count_icd_per_month <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( month ) %>%
  summarise(count_icd = n_distinct(patient_num)) %>%
  mutate( period = ifelse( month < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group_per_month <- count_icd_per_month %>%
  left_join(count_disorder_group_per_month, 
            by = c("month", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_icd

dgPerc_per_month <- perc_disorder_group_per_month %>%  
  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Months",
       y = "Patient (%)",
       title = "Percentage of monthly patients with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_per_month
```

>>>>

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
```

@@

```{r}
# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered_age %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
```

@@

```{r}
# Filtering hospitalizations where patient age is in inclusion criteria
observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

@@

clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
            by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation, age_time_diagnosis)

clinical_table_ado <- clinical_table %>% 
  inner_join(clinical_test,
             by = c("patient_num", "n_hospitalisation"))

@@

all_codes_ado <- all_codes %>% 
   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
```

@@

### Total hospitalize patients age range 11-17

```{r}
unique_patients_hospitalized_per_week <- clinical_table_ado %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( weeks ) %>%
  summarise(count = n_distinct(patient_num))

th_week <- admission_plot_count(df = unique_patients_hospitalized_per_week, 
                     x = "weeks",
                     y = "count", 
                     title = "Total hospitalization (per week)", 
                     y_title = "Patient Count") +
  labs( subtitle = "(age range 11-17)")
th_week
```

## Total number of patients filtering by psychiatric disorders

@@

all_codes_desc <- all_codes_ado %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

@@

### Total patients hospitalized comparing psychiatric disorders vs other

thpd_psyCount <- count_icd_per_week %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by week)",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount

```{r}
thpd_psyPer <- count_icd_per_week %>%
  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  labs(y = "Percentage",
       x = "Date (by week)",
       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer
```

##### Total hospitalizations Per week

@@

patients_hospitalized_per_week_agg_sex_perc %>%
  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage ",
       x = "Week",
       title = "Weekly patient (%) aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```
